---
title: "Homework 6: R Basics"
output: html_document
---

Though not necessary, we can increase the chances
that your script will be reproducible if we set the seed for 
our random number generator.  Leave this code unchanged

```{r setSeed}
set.seed(1)
```


# The binomial distribution

The binomial distribution $\mathrm{Binomial}(n,p)$ is defined by the number of successes in $n$ independent trials, each have probability $p$ of success. Think of flipping a coin $n$ times, where the coin is weighted to have probability $p$ of landing on heads.

The R function 'rbinom' generates random variables with a binomial distribution:  (note that you don't need to name these arguments)

```{r example}
rbinom(n=20, size=10, prob=0.5)
```

produces 20 independent observations from the $\mathrm{Binomial}(10,0.5)$ distribution


# Problem 1 (15 pts)

## Problem 1.1 

Generate 500 random values from the $\mathrm{Binomial}(15,0.5)$ distribution, and store them in a vector called 'binomialDraws' 

Extract and display the first 10 items and then the last 10 items 

```{r q11}
###
binomialDraws = rbinom(500, size = 15, prob = 0.5)
###
binomialDraws[1:10] # First 10 items
tail(binomialDraws, 10) # Last 10 items
```


## Problem 1.2

Find the data type of the items in 'binomialDraws' using 'typeof'

Convert it to a vector of characters, storing this as a new vector 'binomialDrawsChar'

```{r q12}
##
typeof(binomialDraws)
binomialDrawsChar = as.character(binomialDraws)
```
#### Leave this code unchanged:

```{r q12answer}
#typeof(binomialDrawsChar)
```


## Problem 1.3

Generate 500 $\mathrm{Binomial}(15,p)$ for each p ranging over 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, and 0.8, storing the results in vectors called 'binomialDraws2', 'binomialDraws3', ..., 'binomialDraws8'

* Form a matrix called 'binomialDrawsMat' with these 7 vectors (in increasing order, referring to p) as its columns using cbind 

* Report the first 5 rows and then the sub matrix that (has the last 5 rows and last 5 columns).  The nrow/ncol functions may be useful.

* Using 'max', find the maximum value of the second column

```{r q13}
binomialDraws2 = rbinom(500, size = 15, prob = 0.2)
binomialDraws3 = rbinom(500, size = 15, prob = 0.3)
binomialDraws4 = rbinom(500, size = 15, prob = 0.4)
binomialDraws5 = rbinom(500, size = 15, prob = 0.5)
binomialDraws6 = rbinom(500, size = 15, prob = 0.6)
binomialDraws7 = rbinom(500, size = 15, prob = 0.7)
binomialDraws8 = rbinom(500, size = 15, prob = 0.8)
###################
# Create matrix binomialDrawsMat
binomialDrawsMat = cbind(binomialDraws2, binomialDraws3, binomialDraws4, binomialDraws5, binomialDraws6, binomialDraws7, binomialDraws8)
##
# First 5 rows
head(binomialDrawsMat, 5)
# last 5 rows/last 5columns
binomialDrawsMat[(nrow(binomialDrawsMat)-4):nrow(binomialDrawsMat),
                       (ncol(binomialDrawsMat)-4):ncol(binomialDrawsMat)]
# max of second column
max(binomialDrawsMat[, 2])
###
```

## Problem 1.4

Lastly, compute the column (sample) means and subtract this vector from the vector of known expected values of the binomial (that is, $((\textrm{# trials})*p$)

```{r q14}
###
column_means = colMeans(binomialDrawsMat)
###
expected_vals = rep(15 * c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8))
expected_vals - column_means
```

# Problem 2 (15 pts)

## Problem 2.1

Redo problem 1.3 but with 5 million draws from each binomial, accumulating them in a matrix via cbind called 'binomialDrawsBigMat'

Augment 'binomialDrawsBigMat' with the 500 draws from problem 1.3 as new rows.  Report the dimensions of 'binomialDrawsBigMat'

```{r q21}
binomialBigDraws2 = rbinom(5000000, size = 15, prob = 0.2)
binomialBigDraws3 = rbinom(5000000, size = 15, prob = 0.3)
binomialBigDraws4 = rbinom(5000000, size = 15, prob = 0.4)
binomialBigDraws5 = rbinom(5000000, size = 15, prob = 0.5)
binomialBigDraws6 = rbinom(5000000, size = 15, prob = 0.6)
binomialBigDraws7 = rbinom(5000000, size = 15, prob = 0.7)
binomialBigDraws8 = rbinom(5000000, size = 15, prob = 0.8)
################
binomialDrawsBigMat = cbind(binomialBigDraws2, binomialBigDraws3, binomialBigDraws4, binomialBigDraws5, binomialBigDraws6, binomialBigDraws7, binomialBigDraws8)
#
binomialDrawsSmallMat <- cbind(binomialDraws2[1:500], binomialDraws3[1:500],
                               binomialDraws4[1:500], binomialDraws5[1:500],
                               binomialDraws6[1:500], binomialDraws7[1:500],
                               binomialDraws8[1:500])

binomialDrawsBigMat <- rbind(binomialDrawsBigMat, binomialDrawsSmallMat)
##### Repeat question 1.3
# First 5 rows
head(binomialDrawsBigMat, 5)
# last 5 rows/last 5columns
binomialDrawsBigMat[(nrow(binomialDrawsBigMat)-4):nrow(binomialDrawsBigMat),
                       (ncol(binomialDrawsBigMat)-4):ncol(binomialDrawsBigMat)]
# max of second column
max(binomialDrawsBigMat[, 2])
dim(binomialDrawsBigMat)
```

## Problem 2.2

As in problem 1.4,  compute the column (sample) means and subtract this vector from the vector of known expected values of the binomial (that is, $n*p$)

```{r q22}
###
bigmat_colmeans = colMeans(binomialDrawsBigMat)
expected_vals = rep(15 * c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8))
expected_vals - bigmat_colmeans
```

## Problem 3 (10 pts)

An interesting question is how much space do each of the main containers use in R?  Let's compare the amount of memory each of **matrix**, **data.frame**, **list**, and **vector** take.

Create a dataframe called 'binomialDrawsBigDf' from 'binomialDrawsBigMat'. To do this, we can
use 'data.frame' of the matrix object 'binomialDrawsBigMat'.  

Then, using 'list', convert 'binomialDrawsBigDf' into a list, calling it 'binomialDrawsBigList'

Compare the size of each object in bytes by using 'object.size' 

(add to your comparison the atomic vector created by using 'unlist' on binomialDrawsBigList as well)

```{r q3}
###
# 
binomialDrawsBigDf   = data.frame(binomialDrawsBigMat)
binomialDrawsBigList = as.list(binomialDrawsBigDf)
#
format(object.size(binomialDrawsBigMat), unit = 'auto')
format(object.size(binomialDrawsBigDf), unit = 'auto')
format(object.size(binomialDrawsBigList), unit = 'auto')
format(object.size(unlist(binomialDrawsBigList)), unit = 'auto')
###
```
(if you want to make your output in a more 'readable' fashion, you can use format(object.size(...), unit = 'auto'))

Is there any substantial difference in these data structures as to memory usage?

When comparing the matrix, dataframe, and list structures, memory using is essentially the same. However, the unlisted atomic vector uses a considerably greater amount of memory than any of the other three data structures

# Problem 4 (10 pts)

## Problem 4.1

Let's create a 10 x 4 matrix of consecutive integers called 'myMat'.  Make sure that these integers increase consecutively along the rows, not columns

```{r q41}
##
myMat = matrix(1:40, nrow = 10, byrow = T)
###
```

## Problem 4.2

Using logical indexing, modify myMat so that every even number in this matrix is multiplied by  100 every odd number is left alone. This should require just a small bit of code (and the use of the modulus operator \%\% )

```{r q42}
###
#
myMat[myMat %% 2 == 0] = myMat[myMat %% 2 == 0] * 100
###
```

#### Leave this code unchanged:

```{r q42answer}
#myMat
```


## Problem 4.3

Using the 'ifelse' command, return myMat to its original values (note that ifelse vectorizes, try it!)


```{r q43}
###
#
myMat = ifelse(myMat %% 2 == 0, myMat / 100, myMat)
###
```

#### Leave this code unchanged:

```{r q43answer}
#myMat
```
